\thetitle

### 2008-05-29 paper.tex
break line for RREQ/RREP/RERR.
More notation fixup.

\newcommand{\X}{\mathsf{X}}
\newcommand{\M}{\mathsf{M}}
\newcommand{\E}[1]{\left<#1\right>}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Qt}{Q_{\text{time}}}
\newcommand{\Qb}{Q_{\text{byte}}}
\newcommand{\Qi}{Q_{\text{ival}}}
\newcommand{\Di}{Q^{-1}_{\text{ival}}}
\newcommand{\ones}[1]{\mathbf{1}_{#1}}
\newcommand{\zeros}[1]{\mathbf{0}_{#1}}
\newcommand{\mat}[1]{\mathbf{#1}}
\newcommand{\mean}[1]{\bar{#1}}

@@

We represent start time of each flow using an indicator vector. Unlike node indices, however, time is continuous, so we must first map times onto a range of time indices $\set{1,\dots,d_t}$. We transform the range $[0,t_{\max}]$ of possible start times using a simple quantization function:
\begin{align}
\label{eqn:time-quantization}
\Qt(t) &= \max\set{1,\ceil{d_t\fracpx{t}{t_{\max}}}}.
\end{align}
%Suppose the $i$th flow has a quantized start index of $j$ and a end index of $k$. Then $b_{ij'}=\delta{jj'}$ and $e_{ik'}=\delta_{kk'}$.
Let $k$ be the quantized start index of a flow. Then the start vectors for the flow is $\vec{a}=\mathbf{e}_k$ ($\mathbf{e}_k$ has 1 at index $k$ and zeros elsewhere). The advantage of this representation becomes clear when we aggregate flows: the sum of start vectors over a collection of flows is a $d_t$-dimensional histogram indicating the frequency of start times occurring in each quantization bin. In practice, we use $d_t=1500$.

@@

We represent flow size with another indicator vector. Although the set of possible flow sizes is discrete, for practical reasons we map the very large range of possible flow sizes down to a much smaller range of indices (some flows in our trace have 109\caps{MB} of data in a single ten-minute period, which would give an indicator vector of over 100 million dimensions). We use the following function to convert flow sizes to indexes:
\begin{align*}
%\label{eqn:byte-quantization}
\Qb(b) &= \max\set{1,\ceil{d_b\fracp{b-1}{b_{\max}-1}^\beta}}.
\end{align*}
Here $b_{\max}$ is the maximum input flow size. The parameter $\beta$ is a non-linear scaling factor exponent that allows the quantization to smoothly shift from byte-sized resolution at the low end of the possible range of flow sizes to the a resolution of megabytes at the high end of the spectrum. In our example data, $b_{\max}\approx109\text{\caps{MB}}$ as we previously mentioned. We use a representation dimensionality of $d_b=1500$, to match our other quantized vectors, and $\beta=\fracx{1}{2.7}$ because it yields the desired dynamic range response of the possible flow sizes.

@@

The packets size behavior of a flow is represented by a histogram vector, $\vec{z}$, of packet sizes. Explicitly,  $z_{ij}$ is the number of packets in the $i$th flow with $j$ bytes of payload data. The dimension, $d_z$, is the largest possible packet size, which in practice is the maximum transfer unit of the wireless medium, typically 1500 bytes. The representation of inter-packet intervals as a vector, $\vec{v}$, is similar, but the domain of possible intervals is a continuum, so we must map this range onto the discrete index values $\left\{1,\dots,d_v\right\}$. We use similar non-linear quantization function as before:
\begin{align}
\label{eqn:interval-quantization}
\Qi(v) &= \max\set{1,\ceil{d_v\fracp{v-v_{\min}}{v_{\max}-v_{\min}}^\gamma}}.
\end{align}
The element $v_{ij}$ is the number of intervals, $v$, for the $i$th flow such that $\Qi(v)=i$.
The parameter values defining $\Qi$ must be chosen by the experimenter. In our case, the time measurements had microsecond precision, so $v_{\min}=10^{-6}$. The length of each simulated scenario is ten minutes, so the upper limit on intervals is $v_{\max}=600$. The number of dimensions was chosen to match the dimensions for packet size: $d_v=1500$. Finally, the parameter $\gamma\in(0,1)$ allows us to fine-tune how the resolution scales with interval size; we have chosen $\gamma=\frac{1}{3}$ because this choice provides resolution that scales smoothly from the microseconds for small intervals to seconds for large intervals.

@@

The number of flows between each pair of nodes in the network is expressed by the matrix product $\mat{SD}^t\in\R^{n \times n}$ ($*^t$~denotes transpose). That is $(\mat{SD}^t)_{ij}$ is the number of flows from node $i$ to node $j$. The total number of source flows for each node is given by the vector $\mat{S}^t\ones{f}$, and the total number of destination flows is given by $\mat{D}^t\ones{f}$. ($\ones{f}$ is the $f$-dimensional column vector consisting of all ones.)

@@

From the matrices $\mat{Z}$ and $\mat{V}$, we can compute the expected average packet size, $\mean{z}$, and average inter-packet interval, $\mean{v}$, for each flow. The computations require knowing the values that each matrix column represents. The computed duration is only approximate because information is irrecoverably lost in the quantization process (see \Equation{interval-quantization}). The vectors of representative values for $\mat{Z}$ and $\mat{V}$ are:
\begin{align}
\vec{\sigma}&=\seq{1,2,3,\dots,d_z},\\
\vec{\tau}&=\seq{\Di\left(k,\fracx{1}{2}\right)}_{k=1}^{d_v}.
\end{align}
Using these representative values, the expected average packet size for a flow may be computed as $\mean{z}=\fracx{\vec{z}\cdot\vec{\sigma}}{\vec{z}\cdot\ones{d_z}}$, and the expected average inter-packet interval size as $\mean{v}=\fracx{\vec{v}\cdot\vec{\tau}}{\vec{v}\cdot\ones{d_v}}$.

@@

The final form of homogenization we consider, is regularization. This is a restricted form of marginality: we may assume that marginal distributions apply to all the flows in certain groupings. For example, we can provide a distribution of flow sizes on a per-source basis. That is, each source node can have its \textit{own} distribution, and the sizes of its flows are drawn from that distribution, independent of their other properties. This is simply accomplished by left-multiplication by a non-uniform matrix. For example, to make flow behavior uniform on a per source basis, we left-multiply by the ``source-uniformizing'' matrix: $\mat{F}'=(\mat{SS}^t/f)\mat{F}$. Per destination uniformity can be achieved similarly, with left-multiplication by a corresponding destination-uniformizing matrix: $F'=(\mat{DD}^t/f)\mat{F}$.
%Although assumptions about uniformity across flows are common, there is little agreement about which notion of uniformity is most appropriate; often it is simply left unspecified. The confusion arises due to the lack of a unique way of associating flows with nodes. Each node can equally viably be associated with its source or its destination. Which association is more valid? Consider a web request: should the behavior of the request be determined by what the client typically does? Or should the requests that the server typically receives determine behavior? There is no armchair answer: only analysis of real data can give us the answer. The matrix model provides an analytical framework for doing so.
The matrix model has the benefit of making the uniformization of behaviors completely explicit. It also allows us to easily define intermediate choices. If we wish make the behavior of each flow a weighted average of the aggregate behaviors of its source and destination nodes, we can express this easily:
\begin{align}
\mat{F}' = \frac{1}{f}((1-\alpha)(\mat{SS}^t)+\alpha(\mat{DD}^t))\mat{F}.
\end{align}
The parameter $\alpha$ allows us to smoothly vary the weighting of the source and destination behaviors. Source-only behavior is given by $\alpha=0$; destination-only by $\alpha=1$; to weight source and destination equally, take $\alpha=\frac{1}{2}$.

@@

\begin{table*}
\begin{center}
\small
\begin{tabular}{|c|c|l|}
\multicolumn{1}{c}{\textbf{model}} &
\multicolumn{1}{c}{\textbf{transformation}} &
\multicolumn{1}{c}{\textbf{description}} \\
\hline
{\footnotesize{GMM}} & --- & the general matrix model \\\hline\hline
time uniform & $\mat{A}'=\mat{A}(\ones{d_t \times d_t}/d_t)$ & all start times are equally likely \\\hline
node uniform & $\mat{S}'=\mat{S}(\ones{n \times n}/n)$~~~~$\mat{D}'=\mat{D}(\ones{n \times n}/n)$ & nodes behave stochastically identically with respect to flows \\\hline
full uniform & \textit{all uniform transformations} & applies all the uniformity assumptions of the above three models \\\hline\hline
time marginal & $\mat{A}'=(\ones{f \times f}/f)\mat{A}$ & start times chosen from marginal distribution of start times \\\hline
size marginal & $\mat{B}'=(\ones{f \times f}/f)\mat{B}$ & flow size (bytes) is chosen independently of other properties \\\hline
node marginal & $\mat{S}'=(\ones{f \times f}/f)\mat{S}$~~~~$\mat{D}'=(\ones{f \times f}/f)\mat{D}$ & nodes behave distinctly, but behavior is independent of flows \\\hline
packet marginal & $\mat{Z}'=(\ones{f \times f}/f)\mat{Z}$~~~~$\mat{V}'=(\ones{f \times f}/f)\mat{V}$ & packet behavior is variable bit-rate, aggregated across all flows \\\hline
full marginal & \textit{all marginal transformations} & applies all the marginality assumptions of the above four models \\\hline\hline
source regular & $\mat{F}'=(\mat{SS}^t/f)\mat{F}$ & flow behaviors are aggregated on a per source-node basis \\\hline
dest regular & $\mat{F}'=(\mat{DD}^t/f)\mat{F}$ & flow behaviors are aggregated on a per destination-node basis \\\hline
mixed regular & $\mat{F}'=\frac{1}{2f}(\mat{SS}^t+\mat{DD}^t)\mat{F}$ & flow behaviors are aggregated on a mixed source/destination basis \\\hline
\end{tabular}
\caption{Matrix-based traffic models evaluated by paired differential simulation.}
\label{tab:traffic-models}
\end{center}
\vspace{-2em}
\end{table*}

@@

We have selected six performance metrics to present here. They are commonly used as indicators of network performance at the application, network, and link layers of the protocol stack:
\begin{enumerate}
\setlength{\itemsep}{0em}
\item \textbf{Application:} average end-to-end delay, packet delivery ratio, received throughput.
\item \textbf{Network:} \caps{AODV} control overhead (\caps{RREQ/RREP/RERR}), packets dropped in routing queues.
\item \textbf{Link:} 802.11 control overhead (\caps{RTS/CTS/ACK}).
\end{enumerate}
These metrics are commonly used to evaluate wireless protocols. We have examined a broad variety of other wireless network performance metrics, and although we do not have room to present or discuss them here, the results shown are representative of the overall realism of the traffic models.
%For each of the 144 10-minute trace segments, we have run simulations using each of the nine traffic models, for a total of 1,296 simulations.
%Unfortunately, not all of the simulations were able to run to completion. In some scenarios, the number of nodes and flows caused Qualnet to simply fail. We were, however, able to get complete results for 72.5\% of the simulations, providing a clear view of overall trends.\!\footnote{That 27.5\% of simulations did not finish may seem like a cause for concern, but it does not in fact invalidate our results. The simulations that did not finish are those for which the number of flows and nodes overwhelmed the capacity of the simulator. Qualnet cannot simulate more than circa 650 nodes sending 35,000 flows. While it would be of great interest to examine results for these large-scale scenarios, they are essentially useless for comparison to other simulations studies: other studies do not simulate this quantity of nodes and flows precisely because even the best existing simulators do not support it. Once the simulators become more scalable and robust, our analysis can be extended to these scenarios as well.}

@@

\noindent
\textbf{Theorem.} The unique differentiable function satisfying Eqs.~\ref{eqn:normalization-invariance}~\&~\ref{eqn:linear-compounding} is $E(1,e) \ln(y/x)$. \textit{Proof:} Let $f(z) = E(1,z)$. Eq.~\ref{eqn:normalization-invariance} gives $E(x,y) = E(1,y/x) = f(y/x)$. Eq.~\ref{eqn:linear-compounding} gives: $f(z) = f(z/w) + f(w)$. Differentiation by $z$ yields $f'(z) = w^{-1} f'(z/w)$. In particular, if we choose $w=z$, we get $f'(z) = z^{-1} f'(1)$. Integration by $z$ gives:
$f(z) = f'(1) \int z^{-1} dz + c = f'(1) \ln(z) + c$.
By Eq.~\ref{eqn:linear-compounding}, $f(1)=0$, so $c=0$. Thus $f(e)=f'(1)\ln(e)=f'(1)$. We conclude that $f(z)=f(e)\ln(z)$, so $E(x,y) = f(y/x) = f(e)\ln(y/x) = E(1,e)\ln(y/x)$, as desired.\hfill%\QED

>>>>

\newcommand{\X}{\mathsf{X}}
\newcommand{\M}{\mathsf{M}}
\newcommand{\E}[1]{\left<#1\right>}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Qt}{Q_{\text{time}}}
\newcommand{\Qb}{Q_{\text{byte}}}
\newcommand{\Qi}{Q_{\text{ival}}}
\newcommand{\Di}{Q^{-1}_{\text{ival}}}
\newcommand{\ones}[1]{\mathbf{1}_{#1}}
\newcommand{\zeros}[1]{\mathbf{0}_{#1}}
\newcommand{\mat}[1]{\mathbf{#1}}
\newcommand{\mean}[1]{\bar{#1}}
\newcommand{\trans}[1]{{#1}^T}
\newcommand{\inner}[2]{{#1}\cdot{#2}}

@@

We represent start time of each flow using an indicator vector. Unlike node indices, however, time is continuous, so we must first map times onto a range of time indices $\set{1,\dots,d_t}$. We transform the range $[0,t_{\max}]$ of possible start times using a simple quantization function:
\begin{align}
\label{eqn:time-quantization}
\Qt(t) &= \max\set{1,\ceil{d_t\fracpx{t}{t_{\max}}}}.
\end{align}
%Suppose the $i$th flow has a quantized start index of $j$ and a end index of $k$. Then $b_{ij'}=\delta{jj'}$ and $e_{ik'}=\delta_{kk'}$.
Let $k$ be the quantized start index of a flow. Then the start vectors for the flow is $\vec{a}=\vec{e}_k$ (the $k$th standard basis vector, having 1 at index $k$ and zeros elsewhere). The advantage of this representation becomes clear when we aggregate flows: the sum of start vectors over a collection of flows is a $d_t$-dimensional histogram indicating the frequency of start times occurring in each quantization bin. In practice, we use $d_t=1500$.

@@

We represent flow size with another indicator vector. Although the set of possible flow sizes is discrete, for practical reasons we map the very large range of possible flow sizes down to a much smaller range of indices (some flows in our trace have 109\caps{MB} of data in a single ten-minute period, which would give an indicator vector of over 100 million entries). We use the following function to convert flow sizes to indexes:
\begin{align*}
%\label{eqn:byte-quantization}
\Qb(b) &= \max\set{1,\ceil{d_b\fracp{b-1}{b_{\max}-1}^\beta}}.
\end{align*}
Here $b_{\max}$ is the maximum input flow size. The parameter $\beta$ is a non-linear scaling factor exponent that allows the quantization to smoothly shift from byte-sized resolution at the low end of the possible range of flow sizes to the a resolution of megabytes at the high end of the spectrum. In our example data, $b_{\max}\approx109\text{\caps{MB}}$ as we previously mentioned. We use a representation dimensionality of $d_b=1500$, to match our other quantized vectors, and $\beta=\fracx{1}{2.7}$ because it yields the desired dynamic range response of the possible flow sizes.

@@

The packet size behavior of a flow is represented by a histogram vector, $\vec{z}$, of packet sizes. Explicitly,  $z_{ij}$ is the number of packets in the $i$th flow with $j$ bytes of payload data. The dimension, $d_z$, is the largest possible packet size, which in practice is the maximum transfer unit of the wireless medium, typically 1500 bytes. The representation of inter-packet intervals as a vector, $\vec{v}$, is similar, but the domain of possible intervals is a continuum, so we must map this range onto the discrete index values $\left\{1,\dots,d_v\right\}$. We use similar non-linear quantization function as before:
\begin{align}
\label{eqn:interval-quantization}
\Qi(v) &= \max\set{1,\ceil{d_v\fracp{v-v_{\min}}{v_{\max}-v_{\min}}^\gamma}}.
\end{align}
The element $v_{ij}$ is the number of intervals, $v$, for the $i$th flow such that $\Qi(v)=i$.
The parameter values defining $\Qi$ must be chosen by the experimenter. In our case, the time measurements had microsecond precision, so $v_{\min}=10^{-6}$. The length of each simulated scenario is ten minutes, so the upper limit on intervals is $v_{\max}=600$. The number of dimensions was chosen to match the dimensions for packet size: $d_v=1500$. Finally, the parameter $\gamma\in(0,1)$ allows us to fine-tune how the resolution scales with interval size; we have chosen $\gamma=\frac{1}{3}$ because this choice provides resolution that scales smoothly from the microseconds for small intervals to seconds for large intervals.

@@

The number of flows between each pair of nodes in the network is expressed by the matrix product $\trans{\mat{SD}}\in\R^{n \times n}$ ($\trans{*}$~denotes transpose). That is $(\trans{\mat{SD}})_{ij}$ is the number of flows from node $i$ to node $j$. The total number of source flows for each node is given by the vector $\trans{\mat{S}}\ones{f}$, and the total number of destination flows is given by $\trans{\mat{D}}\ones{f}$. ($\ones{f}$ is the $f$-dimensional column vector consisting of all ones.)

@@

From the matrices $\mat{Z}$ and $\mat{V}$, we can compute the expected average packet size, $\mean{z}$, and average inter-packet interval, $\mean{v}$, for each flow. The computations require knowing the values that each matrix column represents. The computed duration is only approximate because information is irrecoverably lost in the quantization process (see \Equation{interval-quantization}). The vectors of representative values for $\mat{Z}$ and $\mat{V}$ are:
\begin{align}
\vec{\sigma}&=\seq{1,2,3,\dots,d_z},\\
\vec{\tau}&=\seq{\Di\left(k,\fracx{1}{2}\right)}_{k=1}^{d_v}.
\end{align}
Using these representative values, the expected average packet size and expected average inter-packet interval size for a flow may be computed as
\begin{align}
  \mean{z}&=\fracx{\inner{\vec{z}}{\vec{\sigma}}}{\inner{\vec{z}}{\ones{d_z}}},\\
  \mean{v}&=\fracx{\inner{\vec{v}}{\vec{\tau}}}{\inner{\vec{v}}{\ones{d_v}}},
\end{align}
respectively, where $\inner{\vec{x}}{\vec{y}}$ denotes the standard inner product of vectors.

@@

The final form of homogenization we consider, is regularization. This is a restricted form of marginality: we may assume that marginal distributions apply to all the flows in certain groupings. For example, we can provide a distribution of flow sizes on a per-source basis. That is, each source node can have its \textit{own} distribution, and the sizes of its flows are drawn from that distribution, independent of their other properties. This is simply accomplished by left-multiplication by a non-uniform matrix. For example, to make flow behavior uniform on a per source basis, we left-multiply by the ``source-uniformizing'' matrix: $\mat{F}'=(\trans{\mat{SS}}/f)\mat{F}$. Per destination uniformity can be achieved similarly, with left-multiplication by a corresponding destination-uniformizing matrix: $F'=(\trans{\mat{DD}}/f)\mat{F}$.
%Although assumptions about uniformity across flows are common, there is little agreement about which notion of uniformity is most appropriate; often it is simply left unspecified. The confusion arises due to the lack of a unique way of associating flows with nodes. Each node can equally viably be associated with its source or its destination. Which association is more valid? Consider a web request: should the behavior of the request be determined by what the client typically does? Or should the requests that the server typically receives determine behavior? There is no armchair answer: only analysis of real data can give us the answer. The matrix model provides an analytical framework for doing so.
The matrix model has the benefit of making the uniformization of behaviors completely explicit. It also allows us to easily define intermediate choices. If we wish make the behavior of each flow a weighted average of the aggregate behaviors of its source and destination nodes, we can express this easily:
\begin{align}
\mat{F}' = \frac{1}{f}((1-\alpha)(\trans{\mat{SS}})+\alpha(\trans{\mat{DD}}))\mat{F}.
\end{align}
The parameter $\alpha$ allows us to smoothly vary the weighting of the source and destination behaviors. Source-only behavior is given by $\alpha=0$; destination-only by $\alpha=1$; to weight source and destination equally, take $\alpha=\frac{1}{2}$.

@@

\begin{table*}
\begin{center}
\small
\begin{tabular}{|c|c|l|}
\multicolumn{1}{c}{\textbf{model}} &
\multicolumn{1}{c}{\textbf{transformation}} &
\multicolumn{1}{c}{\textbf{description}} \\
\hline
{\footnotesize{GMM}} & --- & the general matrix model \\\hline\hline
time uniform & $\mat{A}'=\mat{A}(\ones{d_t \times d_t}/d_t)$ & all start times are equally likely \\\hline
node uniform & $\mat{S}'=\mat{S}(\ones{n \times n}/n)$~~~~$\mat{D}'=\mat{D}(\ones{n \times n}/n)$ & nodes behave stochastically identically with respect to flows \\\hline
full uniform & \textit{all uniform transformations} & applies all the uniformity assumptions of the above three models \\\hline\hline
time marginal & $\mat{A}'=(\ones{f \times f}/f)\mat{A}$ & start times chosen from marginal distribution of start times \\\hline
size marginal & $\mat{B}'=(\ones{f \times f}/f)\mat{B}$ & flow size (bytes) is chosen independently of other properties \\\hline
node marginal & $\mat{S}'=(\ones{f \times f}/f)\mat{S}$~~~~$\mat{D}'=(\ones{f \times f}/f)\mat{D}$ & nodes behave distinctly, but behavior is independent of flows \\\hline
packet marginal & $\mat{Z}'=(\ones{f \times f}/f)\mat{Z}$~~~~$\mat{V}'=(\ones{f \times f}/f)\mat{V}$ & packet behavior is variable bit-rate, aggregated across all flows \\\hline
full marginal & \textit{all marginal transformations} & applies all the marginality assumptions of the above four models \\\hline\hline
source regular & $\mat{F}'=(\trans{\mat{SS}}/f)\mat{F}$ & flow behaviors are aggregated on a per source-node basis \\\hline
dest regular & $\mat{F}'=(\trans{\mat{DD}}/f)\mat{F}$ & flow behaviors are aggregated on a per destination-node basis \\\hline
mixed regular & $\mat{F}'=\frac{1}{2f}(\trans{\mat{SS}}+\trans{\mat{DD}})\mat{F}$ & flow behaviors are aggregated on a mixed source/destination basis \\\hline
\end{tabular}
\caption{Matrix-based traffic models evaluated by paired differential simulation.}
\label{tab:traffic-models}
\end{center}
\vspace{-2em}
\end{table*}

@@

We have selected six performance metrics to present here. They are commonly used as indicators of network performance at the application, network, and link layers of the protocol stack:
\begin{enumerate}
\setlength{\itemsep}{0em}
\item \textbf{Application:} average end-to-end delay, packet delivery ratio, received throughput.
\item \textbf{Network:} \caps{AODV} control overhead (\caps{RREQ/RREP/\\RERR}), packets dropped in routing queues.
\item \textbf{Link:} 802.11 control overhead (\caps{RTS/CTS/ACK}).
\end{enumerate}
These metrics are commonly used to evaluate wireless protocols. We have examined a broad variety of other wireless network performance metrics, and although we do not have room to present or discuss them here, the results shown are representative of the overall realism of the traffic models.
%For each of the 144 10-minute trace segments, we have run simulations using each of the nine traffic models, for a total of 1,296 simulations.
%Unfortunately, not all of the simulations were able to run to completion. In some scenarios, the number of nodes and flows caused Qualnet to simply fail. We were, however, able to get complete results for 72.5\% of the simulations, providing a clear view of overall trends.\!\footnote{That 27.5\% of simulations did not finish may seem like a cause for concern, but it does not in fact invalidate our results. The simulations that did not finish are those for which the number of flows and nodes overwhelmed the capacity of the simulator. Qualnet cannot simulate more than circa 650 nodes sending 35,000 flows. While it would be of great interest to examine results for these large-scale scenarios, they are essentially useless for comparison to other simulations studies: other studies do not simulate this quantity of nodes and flows precisely because even the best existing simulators do not support it. Once the simulators become more scalable and robust, our analysis can be extended to these scenarios as well.}

@@

\noindent
\textbf{Theorem.} The unique differentiable function satisfying Eqs.~\ref{eqn:normalization-invariance} and \ref{eqn:linear-compounding} is $E(1,e) \ln(y/x)$. \textit{Proof:} Let $f(z) = E(1,z)$. Eq.~\ref{eqn:normalization-invariance} gives $E(x,y) = E(1,y/x) = f(y/x)$. Eq.~\ref{eqn:linear-compounding} gives: $f(z) = f(z/w) + f(w)$. Differentiation by $z$ yields $f'(z) = w^{-1} f'(z/w)$. In particular, if we choose $w=z$, we get $f'(z) = z^{-1} f'(1)$. Integration by $z$ gives:
$f(z) = f'(1) \int z^{-1} dz + c = f'(1) \ln(z) + c$.
By Eq.~\ref{eqn:linear-compounding}, $f(1)=0$, so $c=0$. Thus $f(e)=f'(1)\ln(e)=f'(1)$. We conclude that $f(z)=f(e)\ln(z)$, so $E(x,y) = f(y/x) = f(e)\ln(y/x) = E(1,e)\ln(y/x)$, as desired.\hfill%\QED
